---
title: "Figure 3"
output: html_document
date: "2024-11-14"
---

#load libraries
```{r}
library(devtools)
library(ggplot2)
library(ggpubr)
library(damr)
library(zeitgebr)
library(ggetho)
library(sleepr)
library(dplyr)
library(viridis)
library(knitr)
library(afex)
library(lme4)
library(reshape2)
```
#clear workspace
```{r}
rm(list = ls())
```
#check working direcetory
```{r}
getwd()
```
#store path variable for genotype DGRP852
```{r}
DATA_DIR <- "/Users/anabelbyars/Desktop/Riddle Lab/Circadian analysis"
```
#check all files are present
```{r}
list.files(DATA_DIR, pattern="*.txt|*.csv")
```
#define metavariables for the corresponding legend within each genotype
```{r}
data <- fread("metadata.csv")
data
```
#link meta data based on genotype
```{r}
metadata <- link_dam_metadata(data, result_dir = DATA_DIR)
metadata
```
#load into a behavr structure
Exclude mated animals not under "OK" status, find and load matching data
```{r}
dt <- load_dam(metadata[status == "OK"])
summary(dt)
```
#rejoin data and divide by counts
```{r}
dt <- rejoin(dt)
dt[, "nd"] <- dt[, "activity"] / dt[, "counts"]
```
#turn back into behavr table
```{r}
dt <- behavr(dt, metadata [status == "OK"])
```
#make 0 the start 7AM lights on time point
```{r}
dt[,t := t - hours(xmv(baseline_time))]
```
#subset for activity on day 26
```{r}
dt_t1 <- subset(dt, timepoint == "t1")
```
#graph activity on day 26
```{r}
graph_t1 <- ggetho(dt_t1, aes(x = t, y = nd, colour = treatment)) + 
  facet_grid(sex ~ genotype) + 
  labs(y = "Activity per fly", x = "Time (h)") + 
  theme_bw() +
  scale_color_manual(values = c("#0073C2FF", "#EFC000FF", "#BEBEBE"), name = "Treatment") +
  scale_fill_manual(values = c("#0073C2FF", "#EFC000FF", "#BEBEBE"), name = "Treatment") +
  theme(strip.background = element_rect(fill = "white")) + 
  theme(strip.text = element_text(colour = "black")) +
  annotate('rect', xmin = -28800, xmax = 0, ymin = 0, ymax = 12, alpha = 0.5, fill = "darkgrey") +
  annotate('rect', xmin = 43200, xmax = 64800, ymin = 0, ymax = 12, alpha = 0.5, fill = "darkgrey") +
  stat_pop_etho() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = "none")

graph_t1
```
#Run ANOVA with Mauchly's test (tests sphericity)
```{r}
aov_result <- aov_ez("id", "nd", dt_t1_301_F, within = "t", between = "treatment")
summary(aov_result)
```
# Run the Friedman test (treatment x time)
```{r}
data_wide <- dcast(dt_t1_301_F, region_id ~ t + treatment, value.var = "nd")

friedman.test(as.matrix(data_wide[, -1]))
```

