---
title: "Figure 4"
output: html_document
date: "2024-11-14"
---

#load libraries
```{r}
library(devtools)
library(ggplot2)
library(ggpubr)
library(damr)
library(zeitgebr)
library(ggetho)
library(sleepr)
library(dplyr)
library(viridis)
library(knitr)
```

#Clear workspace
```{r}
rm(list = ls())
```

#check working direcetory
```{r}
getwd()
```
#store path variable for genotype DGRP852
```{r}
DATA_DIR <- "/Users/anabelbyars/Desktop/Riddle Lab/Circadian analysis"
```
#check all files are present
```{r}
list.files(DATA_DIR, pattern="*.txt|*.csv")
```

#Define metavariables for the corresponding legend within each genotype
```{r}
data <- fread("metadata.csv")
data
```
#Link meta data based on genotype
```{r}
metadata <- link_dam_metadata(data, result_dir = DATA_DIR)
metadata
```
#Load into a behavr structure
Exclude mated animals not under "OK" status, find and load matching data
```{r}
dt <- load_dam(metadata[status == "OK"])
summary(dt)
```

#Rejoin data and divide by counts
```{r}
dt <- rejoin(dt)
dt[, "nd"] <- dt[, "activity"] / dt[, "counts"]
```
#Turn back into behavr table
```{r}
dt <- behavr(dt, metadata [status == "OK"])
```
#Make 0 the start 7AM lights on time point
```{r}
dt[,t := t - hours(xmv(baseline_time))]
```

#activity v time for morning data
```{r}
data_morning <- fread("morning.csv")
data_afternoon <- fread("afternoon.csv")
data_evening <- fread("evening.csv")
```

#Link and prepare metadata for all 3 times
```{r}
m_metadata <- link_dam_metadata(data_morning, result_dir = DATA_DIR)
m_metadata
dt_m <- load_dam(m_metadata[status == "OK"])
summary(dt_m)
dt_m <- rejoin(dt_m)
dt_m[, "nd"] <- dt_m[, "activity"] / dt_m[, "counts"]
dt_m <- behavr(dt_m, m_metadata [status == "OK"])

a_metadata <- link_dam_metadata(data_afternoon, result_dir = DATA_DIR)
a_metadata
dt_a <- load_dam(a_metadata[status == "OK"])
summary(dt_a)
dt_a <- rejoin(dt_a)
dt_a[, "nd"] <- dt_a[, "activity"] / dt_a[, "counts"]
dt_a <- behavr(dt_a, a_metadata [status == "OK"])

e_metadata <- link_dam_metadata(data_evening, result_dir = DATA_DIR)
e_metadata
dt_e <- load_dam(e_metadata[status == "OK"])
summary(dt_e)
dt_e <- rejoin(dt_e)
dt_e[, "nd"] <- dt_e[, "activity"] / dt_e[, "counts"]
dt_e <- behavr(dt_e, e_metadata [status == "OK"])
```

#activity v time for morning data
```{r}
morning_t1 <- subset(dt_m, timepoint == "t1")

afternoon_t1 <- subset(dt_a, timepoint == "t1")

evening_t1 <- subset(dt_e, timepoint == "t1")
```
#graph data at 3 distinct times of day
```{r}
graph1 <- ggetho(morning_t1, aes(x = t, y = nd, colour = treatment)) + 
  stat_pop_etho() +
  facet_grid(sex ~ genotype) +
  labs(y = "Activity", x = "Time (h)") + 
  theme_bw() +
  scale_color_manual(values = c("#0073C2FF", "#EFC000FF", "#BEBEBE"), name = "Treatment") +
  scale_fill_manual(values = c("#0073C2FF", "#EFC000FF", "#BEBEBE"), name = "Treatment") +
  scale_y_continuous(limits = c(0, 15),
                     breaks = seq(0, 15, by = 3)) +
  theme(strip.background = element_rect(fill = "white"), 
        strip.text = element_text(colour = "black"),
        panel.grid.major = element_blank()) 

graph1
```
```{r}
graph2 <- ggetho(afternoon_t1, aes(x = t, y = nd, colour = treatment)) + 
  stat_pop_etho() +
  facet_grid(sex ~ genotype) +
  labs(y = "Activity", x = "Time (h)") + 
  theme_bw() +
  scale_color_manual(values = c("#0073C2FF", "#EFC000FF", "#BEBEBE"), name = "Treatment") +
  scale_fill_manual(values = c("#0073C2FF", "#EFC000FF", "#BEBEBE"), name = "Treatment") +
  scale_y_continuous(limits = c(0, 15),
                     breaks = seq(0, 15, by = 3)) +
  theme(strip.background = element_rect(fill = "white"), 
        strip.text = element_text(colour = "black"),
        panel.grid.major = element_blank())

graph2
```
```{r}
graph3 <- ggetho(evening_t1, aes(x = t, y = nd, colour = treatment)) + 
  stat_pop_etho() +
  facet_grid(sex ~ genotype) +
  labs(y = "Activity", x = "Time (h)") + 
  theme_bw() +
  scale_color_manual(values = c("#0073C2FF", "#EFC000FF", "#BEBEBE"), name = "Treatment") +
  scale_fill_manual(values = c("#0073C2FF", "#EFC000FF", "#BEBEBE"), name = "Treatment") +
  scale_y_continuous(limits = c(0, 15),
                     breaks = seq(0, 15, by = 3)) +
  theme(strip.background = element_rect(fill = "white"), 
        strip.text = element_text(colour = "black"),
        panel.grid.major = element_blank()) 

graph3
```

#Test normality with Kolmogorov-Smirnov test (>5000 values)

```{r}
ks_test_result1 <- ks.test(dt_m$nd, "pnorm", mean = mean(dt_m$nd), sd = sd(dt_m$nd))
print(ks_test_result1)

ks_test_result2 <- ks.test(dt_a$nd, "pnorm", mean = mean(dt_a$nd), sd = sd(dt_a$nd))
print(ks_test_result2)

ks_test_result3 <- ks.test(dt_e$nd, "pnorm", mean = mean(dt_e$nd), sd = sd(dt_e$nd))
print(ks_test_result3)
```

#Ranked ANOVA
```{r}
dt_m$genotype <- as.factor(dt_m$genotype)
dt_m$sex <- as.factor(dt_m$sex)
dt_m$treatment <- as.factor(dt_m$treatment)
dt_m$timepoint <- as.factor(dt_m$timepoint)

a_morning <- dt_m[order(dt_m$nd), ]
a_morning$rank <- 1:nrow(a_morning)
a_morning <- aov(rank ~ genotype * sex * treatment, data = a_morning)
summary(a_morning)
```
```{r}
dt_a$genotype <- as.factor(dt_a$genotype)
dt_a$sex <- as.factor(dt_a$sex)
dt_a$treatment <- as.factor(dt_a$treatment)
dt_a$timepoint <- as.factor(dt_a$timepoint)

a_afternoon <- dt_a[order(dt_a$nd), ]
a_afternoon$rank <- 1:nrow(a_afternoon)
a_afternoon <- aov(rank ~ genotype * sex * treatment, data = a_afternoon)
summary(a_afternoon)
```
```{r}
dt_e$genotype <- as.factor(dt_e$genotype)
dt_e$sex <- as.factor(dt_e$sex)
dt_e$treatment <- as.factor(dt_e$treatment)
dt_e$timepoint <- as.factor(dt_e$timepoint)

a_evening <- dt_e[order(dt_e$nd), ]
a_evening$rank <- 1:nrow(a_evening)
a_evening <- aov(rank ~ genotype * sex * treatment, data = a_evening)
summary(a_evening)
```

#Tukey test
```{r}
t_morning <- TukeyHSD(a_morning, conf.level=0.95)
t_morning
```
```{r}
t_afernoon <- TukeyHSD(a_afternoon, conf.level=0.95)
t_afernoon
```
```{r}
t_evening <- TukeyHSD(a_evening, conf.level=0.95)
t_evening
```


