---
title: "Figure 5"
output: html_document
date: "2024-11-14"
---

#load libraries
```{r}
library(readxl)
library(writexl)
library(survival)
library(survminer)
library(ggplot2)
library(ggfortify)
library(viridis)
library(dplyr)
library(here)
library(ARTool)
```

#Clear workspace
```{r}
rm(list = ls())
```


#check working directory
```{r}
getwd()
```

#Lifespan assay for 5D and 20D treatments combined
set file name and load data for genotype (one at a time)
```{r}
dgrp852 <- read_excel("Longevity_852_2023_04_17.xlsx", sheet = "Master")

dgrp304 <- read_excel("DGRP304_2023_09_25.xlsx", sheet = "Master")

dgrp786 <- read_excel("DGRP786_2023_11_06.xlsx", sheet = "Master")

dgrp301 <- read_excel("DGRP301_2024_06_17.xlsx", sheet = "Master")
```

#make sure parameters are factors so TukeyHSD works
```{r}
dgrp852$sex <- as.factor(dgrp852$sex)
dgrp852$treatment <- as.factor(dgrp852$treatment)

dgrp304$sex <- as.factor(dgrp304$sex)
dgrp304$treatment <- as.factor(dgrp304$treatment)

dgrp786$sex <- as.factor(dgrp786$sex)
dgrp786$treatment <- as.factor(dgrp786$treatment)

dgrp301$sex <- as.factor(dgrp301$sex)
dgrp301$treatment <- as.factor(dgrp301$treatment)
```

#combine data for all genotypes
```{r}
survival_master <- rbind(dgrp304, dgrp786, dgrp852, dgrp301)
```

#sanity check 1
```{r}
km.model_genotype <- survfit(formula = Surv(death_day, status) ~ genotype, data = survival_master)
km.model_genotype


km.model_all <- survfit(formula = Surv(death_day, status) ~ genotype + sex + treatment, data = survival_master)
km.model_all

```

#survival curve with all genotypes censored
```{r}
survival_plot <- ggsurvplot(
  survfit(Surv(death_day, status) ~ treatment, data = survival_master),
  facet.by = c("sex", "genotype"),
  legend.title = "Treatment",
  legend.labs = c("Control", "5-day", "20-day"),
  xlab = "Days",
  palette = c("#BEBEBE", "#EFC000FF", "#0073C2FF"),
  surv.median.line = "v",
  pval = TRUE)

survival_plot
```


#Censor lifespan data to retrieve only status of 1
```{r}
dgrp301_cen <- subset(dgrp301, status == 1)
dgrp304_cen <- subset(dgrp304, status == 1)
dgrp786_cen <- subset(dgrp786, status == 1)
dgrp852_cen <- subset(dgrp852, status == 1)
```

#Normality test Shapiro-Wilk
```{r}
shapiro.test(dgrp301_cen$death_day)
shapiro.test(dgrp304_cen$death_day)
shapiro.test(dgrp786_cen$death_day)
shapiro.test(dgrp852_cen$death_day)
```

#Kruskal test for non-normal data
```{r}
kruskal.test(death_day ~ treatment, data = dgrp301_cen)
kruskal.test(death_day ~ sex, data = dgrp301_cen)

```

#ANOVAs (can't use, data not normal)
```
a_dgrp301 <- aov(death_day ~ sex * treatment, 
         data = dgrp301_cen)
summary(a_dgrp301)

a_dgrp304 <- aov(death_day ~ sex * treatment, 
         data = dgrp304_cen)
summary(a_dgrp304)

a_dgrp786 <- aov(death_day ~ sex * treatment, 
         data = dgrp786_cen)
summary(a_dgrp786)

a_dgrp852 <- aov(death_day ~ sex * treatment, 
         data = dgrp852_cen)
summary(a_dgrp852)

```

#Ranked ANOVA for non-normal data
```{r}
a2_301 <- dgrp301_cen[order(dgrp301_cen$death_day), ]
a2_301$rank <- 1:nrow(a2_301)
a2_301 <- aov(rank ~ sex * treatment, data = a2_301)
summary(a2_301)
```
```{r}
a2_304 <- dgrp304_cen[order(dgrp304_cen$death_day), ]
a2_304$rank <- 1:nrow(a2_304)
a2_304 <- aov(rank ~ sex * treatment, data = a2_304)
summary(a2_304)
```
```{r}
a2_786 <- dgrp786_cen[order(dgrp786_cen$death_day), ]
a2_786$rank <- 1:nrow(a2_786)
a2_786 <- aov(rank ~ sex * treatment, data = a2_786)
summary(a2_786)
```
```{r}
a2_852 <- dgrp852_cen[order(dgrp852_cen$death_day), ]
a2_852$rank <- 1:nrow(a2_852)
a2_852 <- aov(rank ~ sex * treatment, data = a2_852)
summary(a2_852)
```

#Tukey tests for comparisons
```{r}
t1_301 <- TukeyHSD(a2_301, conf.level = 0.95)
t1_301
```
```{r}
t1_304 <- TukeyHSD(a2_304, conf.level = 0.95)
t1_304
```
```{r}
t1_786 <- TukeyHSD(a2_786, conf.level = 0.95)
t1_786
```
```{r}
t1_852 <- TukeyHSD(a2_852, conf.level = 0.95)
t1_852
```

---
Median lifespans separated by sex, genotype, and treatment
---
#Fit the survival mode
```{r}
fit <- survfit(Surv(death_day, status) ~ sex + genotype + treatment, data = survival_master)
```
#Get summary
```{r}
fit_summary <- summary(fit)
```
#Extract median survival times
```{r}
median_survival <- fit_summary$table[, "median"]
```
#Create data frame for values
```{r}
median_survival_df <- data.frame(
  Combination = names(median_survival),
  Median_Survival_Days = median_survival)
```
#Print values
```{r}
median_survival_df
```