---
title: "Figure 2"
output: html_document
date: "2024-11-14"
---

#Load the library
```{r}
library(readxl)
library(reshape2)
library(ggplot2)
library(reshape)
library(dplyr)
library(here)
library(ARTool)
```

#Clear workspace
```{r}
rm(list = ls())
```
#check working directory
```{r}
getwd()
```
#DGRP786 genotype data to run:
loading data frames for DAM data
set file name
```{r}
DAM_786 <- "DAM_786_2023_11_06.xlsx"
```
#Steps to load and organize all the data for the t0 timepoint
load the normalized data into a dataframe for each timepoint/treatment
```{r}
  t0_c_786 <- read_excel(DAM_786, sheet = "t0_c_norm")
  t0_5d_786 <- read_excel(DAM_786, sheet = "t0_5d_norm")
```
#load the legend specific to the timepoint t0
```{r}
  legend_786 <- read_excel(DAM_786, sheet = "legend_t0")
  legend_786$chan_no <- as.factor(legend_786$chan_no)
```
#break the legend out by treatment in order to do a left join 
```{r}
  legend_c_786 <- subset(legend_786, treatment == "c")
  legend_5d_786 <- subset(legend_786, treatment== "5d")
```
#convert to data frame
```{r}
  t0_c_786 <- as.data.frame(t0_c_786)
  t0_5d_786 <- as.data.frame(t0_5d_786)
```
#use the melt function to convert to long format
```{r}
  t0_c_786 <- melt(t0_c_786, id.vars = c("index", "date", "time"), 
                   measure.vars = c("1", "2", "3", "4", "5", "6", "7", "8",
                                    "9", "10", "11", "12", "13", "14", "15",
                                    "16", "17", "18", "19", "20", "21", "22",
                                    "23", "24", "25", "26", "27", "28", "29",
                                    "30", "31", "32"))
  t0_5d_786 <- melt(t0_5d_786, id.vars = c("index", "date", "time"), 
                    measure.vars = c("1", "2", "3", "4", "5", "6", "7", "8",
                                     "9", "10", "11", "12", "13", "14", "15",
                                     "16", "17", "18", "19", "20", "21", "22",
                                     "23", "24", "25", "26", "27", "28", "29",
                                     "30", "31", "32"))
```
#sanity check 1
```{r}
  summary(t0_c_786)
  summary(t0_5d_786)
```
#after data is in long format, sum the 5-minute intervals into one daily total
```{r}
  t0_c_786 <- aggregate(t0_c_786$value, by=list(chan_no=t0_c_786$variable), FUN=sum)
  t0_5d_786 <- aggregate(t0_5d_786$value, by=list(chan_no=t0_5d_786$variable), FUN=sum)
```
#rename columns from variable and value
```{r}
  colnames(t0_c_786) <- c("chan_no", "daily_activity")
  colnames(t0_5d_786) <- c("chan_no", "daily_activity")
```
#unblind using the legends- specific to treatment- on the vial_no value
```{r}
  t0_c_786 <- left_join(t0_c_786, legend_c_786, by = "chan_no")
  t0_5d_786 <- left_join(t0_5d_786, legend_5d_786, by = "chan_no")
```
#get rid of unneeded columns
```{r}
  keeps <- c("chan_no", "daily_activity", "treatment", "genotype", "sex", "age")
  t0_c_786 <- t0_c_786[keeps]
  t0_5d_786 <- t0_5d_786[keeps]
```
#combine the treatments into one master per timepoint using rbind
```{r}
  master_t0_786 <- rbind(t0_c_786, t0_5d_786)
```
#sanity check 2
```{r}
  summary(master_t0_786)

  hist(master_t0_786$daily_activity)
  abline(v = mean(master_t0_786$daily_activity), col="red", lwd=3)
```


---
#Steps to load and organize all the data for the t1 timepoint
load the normalized data into a dataframe for each timepoint/treatment
```{r}
  t1_c_786 <- read_excel(DAM_786, sheet = "t1_c_norm")
  t1_5d_786 <- read_excel(DAM_786, sheet = "t1_5d_norm")
  t1_20d_786 <- read_excel(DAM_786, sheet = "t1_20d_norm")
```
#load the legend specific to the timepoint
```{r}
  legend_786 <- read_excel(DAM_786, sheet = "legend_t1")
  legend_786$chan_no <- as.factor(legend_786$chan_no)
```
#break the legend out by treatment in order to do a left join
```{r}
  legend_c_786 <- subset(legend_786, treatment == "c")
  legend_5d_786 <- subset(legend_786, treatment== "5d")
  legend_20d_786 <- subset(legend_786, treatment== "20d")
```
#convert to data frame
```{r}
  t1_c_786 <- as.data.frame(t1_c_786)
  t1_5d_786 <- as.data.frame(t1_5d_786)
  t1_20d_786 <- as.data.frame(t1_20d_786)
```
#use the melt function to convert to long format
```{r}
  t1_c_786 <- melt(t1_c_786, id.vars = c("index", "date", "time"), 
                   measure.vars = c("1", "2", "3", "4", "5", "6", "7", "8",
                                    "9", "10", "11", "12", "13", "14", "15",
                                    "16", "17", "18", "19", "20", "21", "22",
                                    "23", "24", "25", "26", "27", "28", "29",
                                    "30", "31", "32"))
  t1_5d_786 <- melt(t1_5d_786, id.vars = c("index", "date", "time"), 
                    measure.vars = c("1", "2", "3", "4", "5", "6", "7", "8",
                                     "9", "10", "11", "12", "13", "14", "15",
                                     "16", "17", "18", "19", "20", "21", "22",
                                     "23", "24", "25", "26", "27", "28", "29",
                                     "30", "31", "32"))
  t1_20d_786 <- melt(t1_20d_786, id.vars = c("index", "date", "time"), 
                     measure.vars = c("1", "2", "3", "4", "5", "6", "7", "8",
                                      "9", "11", "12", "13", "14", "15",
                                      "16", "17", "18", "19", "20", "21", "22",
                                      "23", "24", "25", "26", "27", "28", "29",
                                      "30", "31", "32"))
```
#sanity check 1
```{r}
 summary(t1_c_786)
  summary(t1_5d_786)
  summary(t1_20d_786)
```
#after data is in long format, sum the 5-minute intervals into one daily total
```{r}
  t1_c_786 <- aggregate(t1_c_786$value, by=list(chan_no=t1_c_786$variable), FUN=sum)
  t1_5d_786 <- aggregate(t1_5d_786$value, by=list(chan_no=t1_5d_786$variable), FUN=sum)
  t1_20d_786 <- aggregate(t1_20d_786$value, by=list(chan_no=t1_20d_786$variable), FUN=sum)
```
#rename columns from variable and value
```{r}
  colnames(t1_c_786) <- c("chan_no", "daily_activity")
  colnames(t1_5d_786) <- c("chan_no", "daily_activity")
  colnames(t1_20d_786) <- c("chan_no", "daily_activity")
```
#unblind using the legends- specific to treatment- on the vial_no value
```{r}
  t1_c_786 <- left_join(t1_c_786, legend_c_786, by = "chan_no")
  t1_5d_786 <- left_join(t1_5d_786, legend_5d_786, by = "chan_no")
  t1_20d_786 <- left_join(t1_20d_786, legend_20d_786, by = "chan_no")
```
#get rid of unneeded columns
```{r}
  keeps <- c("chan_no", "daily_activity", "treatment", "genotype", "sex", "age")
  t1_c_786 <- t1_c_786[keeps]
  t1_5d_786 <- t1_5d_786[keeps]
  t1_20d_786 <- t1_20d_786[keeps]
```
#combine the treatments into one master per timepoint using rbind
```{r}
  master_t1_786 <- rbind(t1_c_786, t1_5d_786, t1_20d_786)
```
#sanity check 2
```{r}
  summary(master_t1_786)

  hist(master_t1_786$daily_activity)
  abline(v = mean(master_t1_786$daily_activity), col="red", lwd=3)
```


---
#Steps to load and organize all the data for the t2 timepoint
load the normalized data into a dataframe for each timepoint/treatment
```{r}
  t2_c_786 <- read_excel(DAM_786, sheet = "t2_c_norm")
  t2_5d_786 <- read_excel(DAM_786, sheet = "t2_5d_norm")
  t2_20d_786 <- read_excel(DAM_786, sheet = "t2_20d_norm")
```
#load the legend specific to the timepoint t1_t2
```{r}
  legend_t1_t2_786 <- read_excel(DAM_786, sheet = "legend_t2")
  legend_t1_t2_786$chan_no <- as.factor(legend_t1_t2_786$chan_no)
```
#break the legend out by treatment in order to do a left join
```{r}
  legend_c_786 <- subset(legend_t1_t2_786, treatment == "c")
  legend_5d_786 <- subset(legend_t1_t2_786, treatment== "5d")
  legend_20d_786 <- subset(legend_t1_t2_786, treatment== "20d")
```
#convert to data frame
```{r}
  t2_c_786 <- as.data.frame(t2_c_786)
  t2_5d_786 <- as.data.frame(t2_5d_786)
  t2_20d_786 <- as.data.frame(t2_20d_786)
```
#use the melt function to convert to long format
```{r}
  t2_c_786 <- melt(t2_c_786, id.vars = c("index", "date", "time"), 
                   measure.vars = c("1", "2", "3", "4", "5", "6", "7", "8",
                                    "9", "10", "11", "12", "13", "14", "15",
                                    "16", "17", "18", "19", "20", "21", "22",
                                    "23", "24", "25", "26", "27", "28", "29",
                                    "30", "31", "32"))
  t2_5d_786 <- melt(t2_5d_786, id.vars = c("index", "date", "time"), 
                    measure.vars = c("1", "2", "3", "4", "5", "6", "7", "8",
                                     "9", "10", "11", "12", "13", "14", "15",
                                     "16", "17", "18", "19", "20", "21", "22",
                                     "23", "24", "25", "26", "27", "28", "29",
                                     "30", "31", "32"))
  t2_20d_786 <- melt(t2_20d_786, id.vars = c("index", "date", "time"), 
                     measure.vars = c("1", "2", "3", "4", "5", "6", "7", "8",
                                      "9", "11", "12", "13", "14", "15",
                                      "16", "17", "18", "19", "20", "21", "22",
                                      "23", "24", "25", "26", "27", "28", "29",
                                      "30", "31", "32"))
```
#sanity check 1
```{r}
  summary(t2_c_786)
  summary(t2_5d_786)
  summary(t2_20d_786)
```
#after data is in long format, sum the 5-minute intervals into one daily total
```{r}
  t2_c_786 <- aggregate(t2_c_786$value, by=list(chan_no=t2_c_786$variable), FUN=sum)
  t2_5d_786 <- aggregate(t2_5d_786$value, by=list(chan_no=t2_5d_786$variable), FUN=sum)
  t2_20d_786 <- aggregate(t2_20d_786$value, by=list(chan_no=t2_20d_786$variable), FUN=sum)
```
#rename columns from variable and value
```{r}
  colnames(t2_c_786) <- c("chan_no", "daily_activity")
  colnames(t2_5d_786) <- c("chan_no", "daily_activity")
  colnames(t2_20d_786) <- c("chan_no", "daily_activity")
```
#unblind using the legends- specific to treatment- on the vial_no value
```{r}
  t2_c_786 <- left_join(t2_c_786, legend_c_786, by = "chan_no")
  t2_5d_786 <- left_join(t2_5d_786, legend_5d_786, by = "chan_no")
  t2_20d_786 <- left_join(t2_20d_786, legend_20d_786, by = "chan_no")
```
#get rid of unneeded columns
```{r}
  keeps <- c("chan_no", "daily_activity", "treatment", "genotype", "sex", "age")
  t2_c_786 <- t2_c_786[keeps]
  t2_5d_786 <- t2_5d_786[keeps]
  t2_20d_786 <- t2_20d_786[keeps]
```
#combine the treatments into one master per timepoint using rbind
```{r}
  master_t2_786 <- rbind(t2_c_786, t2_5d_786, t2_20d_786)
```
#sanity check 2
```{r}
  summary(master_t2_786)

  hist(master_t2_786$daily_activity)
  abline(v = mean(master_t2_786$daily_activity), col="red", lwd=3)
```


---
#Steps to load and organize all the data for the t3 timepoint
load the normalized data into a dataframe for each timepoint/treatment
```{r}
  t3_c_786 <- read_excel(DAM_786, sheet = "t3_c_norm")
  t3_5d_786 <- read_excel(DAM_786, sheet = "t3_5d_norm")
  t3_20d_786 <- read_excel(DAM_786, sheet = "t3_20d_norm")
```
#load the legend specific to the timepoint t3
```{r}
  legend_t3_786 <- read_excel(DAM_786, sheet = "legend_t3")
  legend_t3_786$chan_no <- as.factor(legend_t3_786$chan_no)
```
#break the legend out by treatment in order to do a left join
```{r}
  legend_c_786 <- subset(legend_t3_786, treatment == "c")
  legend_5d_786 <- subset(legend_t3_786, treatment== "5d")
  legend_20d_786 <- subset(legend_t3_786, treatment== "20d")
```
#convert to data frame
```{r}
  t3_c_786 <- as.data.frame(t3_c_786)
  t3_5d_786 <- as.data.frame(t3_5d_786)
  t3_20d_786 <- as.data.frame(t3_20d_786)
```
#use the melt function to convert to long format
```{r}
  t3_c_786 <- melt(t3_c_786, id.vars = c("index", "date", "time"), 
                   measure.vars = c("1", "2", "3", "4", "5", "6", "7", "8",
                                    "9", "10", "11", "12", "13", "14", "15",
                                    "16", "17", "18", "19", "20", "21", "22",
                                    "23", "24", "25", "26", "27", "28", "29",
                                    "30", "31", "32"))
  t3_5d_786 <- melt(t3_5d_786, id.vars = c("index", "date", "time"), 
                    measure.vars = c("1", "2", "3", "4", "5", "6", "7", "8",
                                     "9", "10", "11", "12", "13", "14", "15",
                                     "16", "17", "18", "19", "20", "21", "22",
                                     "23", "24", "25", "26", "27", "28", "29",
                                     "30", "31", "32"))
  t3_20d_786 <- melt(t3_20d_786, id.vars = c("index", "date", "time"), 
                     measure.vars = c("1", "2", "3", "4", "5", "6", "7", "8",
                                      "9", "11", "12", "13", "14", "15",
                                      "16", "17", "18", "19", "20", "21", "22",
                                      "23", "24", "25", "26", "27", "28", "29",
                                      "30", "31", "32"))
```
#sanity check 1
```{r}
  summary(t3_c_786)
  summary(t3_5d_786)
  summary(t3_20d_786)
```
#after data is in long format, sum the 5-minute intervals into one daily total
```{r}
  t3_c_786 <- aggregate(t3_c_786$value, by=list(chan_no=t3_c_786$variable), FUN=sum)
  t3_5d_786 <- aggregate(t3_5d_786$value, by=list(chan_no=t3_5d_786$variable), FUN=sum)
  t3_20d_786 <- aggregate(t3_20d_786$value, by=list(chan_no=t3_20d_786$variable), FUN=sum)
```
#rename columns from variable and value
```{r}
  colnames(t3_c_786) <- c("chan_no", "daily_activity")
  colnames(t3_5d_786) <- c("chan_no", "daily_activity") 
  colnames(t3_20d_786) <- c("chan_no", "daily_activity")
```
#unblind using the legends- specific to treatment- on the vial_no value
```{r}
  t3_c_786 <- left_join(t3_c_786, legend_c_786, by = "chan_no")
  t3_5d_786 <- left_join(t3_5d_786, legend_5d_786, by = "chan_no")
  t3_20d_786 <- left_join(t3_20d_786, legend_20d_786, by = "chan_no")
```
#get rid of unneeded columns
```{r}
  keeps <- c("chan_no", "daily_activity", "treatment", "genotype", "sex", "age")
  t3_c_786 <- t3_c_786[keeps]
  t3_5d_786 <- t3_5d_786[keeps]
  t3_20d_786 <- t3_20d_786[keeps]
```
#combine the treatments into one master per timepoint using rbind
```{r}
  master_t3_786 <- rbind(t3_c_786, t3_5d_786, t3_20d_786)
```
#sanity check 2
```{r}
  summary(t3_c_786)

  hist(master_t3_786$daily_activity)
  abline(v = mean(master_t3_786$daily_activity), col="red", lwd=3)
```
#rbind to master
```{r}
master_786 <- rbind(master_t0_786, master_t1_786, master_t2_786, master_t3_786)
```
#plot for sig 786 (t0,t1,t2)
```{r}

boxplot_786 <- ggplot(data = master_786, aes(x = sex, y = daily_activity)) +
  geom_boxplot(aes(fill = "white", color = treatment), outlier.shape = NA, size=0.75) +
  geom_jitter(aes(color = treatment), width = 0.2, size = 1, alpha = 0.6) +
  facet_grid(~age) +
  scale_fill_manual(values = "white") + 
  scale_color_manual(values = c("#EFC000FF", "#BEBEBE", "#0073C2FF"),
                     labels = c("5-day", "Control", "20-day")) +
  labs(x = "Sex", y = "Activity per fly", fill = "Treatment", color = "Treatment") +
  theme_bw() +
  theme(strip.background = element_rect(fill = "white"), 
        strip.text = element_text(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) 

boxplot_786
```
---
Analysis
---
##Normal distribution Shapiro-Wilk tests
```{r}
shapiro.test(master_t0_786$daily_activity)
shapiro.test(master_t1_786$daily_activity)
shapiro.test(master_t2_786$daily_activity)
shapiro.test(master_t3_786$daily_activity)
```
##Ranked ANOVA for non-normal data, DGRP786 
```{r}
master_t0_786_a <- master_t0_786[order(master_t0_786$daily_activity), ]
master_t0_786_a$rank <- 1:nrow(master_t0_786_a)
master_t0_786_a <- aov(daily_activity ~ sex * treatment, data = master_t0_786)
summary(master_t0_786_a)
```
```{r}
master_t1_786_a <- master_t1_786[order(master_t1_786$daily_activity), ]
master_t1_786_a$rank <- 1:nrow(master_t1_786_a)
master_t1_786_a <- aov(daily_activity ~ sex * treatment, data = master_t1_786)
summary(master_t1_786_a)
```
```{r}
master_t2_786_a <- master_t1_786[order(master_t2_786$daily_activity), ]
master_t2_786_a$rank <- 1:nrow(master_t2_786_a)
master_t2_786_a <- aov(daily_activity ~ sex * treatment, data = master_t2_786)
summary(master_t2_786_a)
```
```{r}
master_t3_786_a <- master_t1_786[order(master_t3_786$daily_activity), ]
master_t3_786_a$rank <- 1:nrow(master_t3_786_a)
master_t3_786_a <- aov(daily_activity ~ sex * treatment, data = master_t3_786)
summary(master_t3_786_a)
```
#Run the Tukey HSD, change the data name
```{r}
TukeyHSD(master_t0_786_a, conf.level = 0.95)
TK <- TukeyHSD(master_t0_786_a)
TK_data<-as.data.frame(TK[3])
```