---
title: "Figure 2"
output: html_document
date: "2024-11-14"
---

#load libraries
```{r}
library(ggplot2)
library(damr)
library(ggetho)
library(dplyr)
library(here)
library(ARTool)
library(rstatix)
```
With "here" we do not need to set a working directory. Keep script and data in same folder as the Data subfolder (has .txt files) and the metadata.csv - It should already be setup and ready to use in the Data 1.zip.

#clear workspace
```{r}
rm(list = ls())
```
#check working direcetory
```{r}
getwd()
```
#set file name
```{r}
DATA_DIR <- "Data"
```
#check all files are present
```{r}
list.files(DATA_DIR, pattern="*.txt|*.csv")
```


#define metavariables for the corresponding legend within each genotype
```{r}
data <- fread("metadata.csv")
data
```
#link metadata based on genotype
```{r}
metadata <- link_dam_metadata(data, result_dir = DATA_DIR)
metadata
```
#load into a behavr structure
Exclude mated animals not under "OK" status, find and load matching data
```{r}
dt <- load_dam(metadata[status == "OK"])
```
#sanity check
```{r}
summary(dt)
```
#rejoin data and divide by counts, nd is normalized activity per fly
```{r}
dt <- rejoin(dt)
dt[, "nd"] <- dt[, "activity"] / dt[, "counts"]
```
#turn back into behavr table
```{r}
dt <- behavr(dt, metadata [status == "OK"])
```
#make lights on at 7AM the 0
```{r}
dt[,t := t - hours(xmv(baseline_time))]
```
#subset data for control groups on day 26
```{r}
genotype_controls1 <- subset(dt, timepoint == "t1")
genotype_controls <- subset(genotype_controls1, treatment == "c")
```

# Shift the time for genotypes 786 and 304 by -3600 seconds
DGRP786 and DGRP304 were the only two genotypes impacted by daylights savings. So set them back by 1hr. In the metadata sheet I already set the 0 with the daylights savings accounted for so instead of 7=0 it was 6=0.

When plotting all of the data together we had to shift 786 and 304 x-axis left 1hr to align with the other axes.
```{r}
genotype_controls$t <- ifelse(genotype_controls$genotype %in% c(786, 304), genotype_controls$t - 3600, genotype_controls$t)
```


#plot the activity for controls on day 26 over 24 hours
```{r}
graph_controls <- ggetho(genotype_controls, aes(x = t, y = nd, colour = treatment)) + 
  stat_pop_etho() + 
  facet_grid(sex ~ genotype) +
  labs(y = "Activity per fly", x = "Time (h)") + 
  theme_bw() +
  scale_color_manual(values = c("purple"), name = "Treatment") +
  scale_fill_manual(values = c("purple"), name = "Treatment") +
  theme(strip.background = element_rect(fill = "white"), 
        strip.text = element_text(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = "none") +
  annotate('rect', xmin = -25200, xmax = 0, ymin = 0, ymax = 12, alpha = 0.3, fill = "darkgrey") +
  annotate('rect', xmin = 43200, xmax = 59500, ymin = 0, ymax = 12, alpha = 0.3, fill = "darkgrey")

graph_controls

```


---
Analysis
---
#aggregate the data by region_id to analyze total activity
```{r}
activity_summary <- genotype_controls %>% 
  group_by(genotype, sex, region_id) %>% 
  summarise(
    mean_act = mean(nd, na.rm = TRUE),
    total_act = sum(nd, na.rm = TRUE),
    .groups = "drop"
  )
```
#average the data for each sex and genotype combination
Values utilized in the manuscript
```{r}
activity_summary_avg <- activity_summary %>%
  group_by(sex, genotype) %>%
  summarise(
    avg_mean_act = mean(mean_act, na.rm = TRUE),
    avg_total_act = mean(total_act, na.rm = TRUE),
    .groups = "drop"
  )

activity_summary_avg
```

#shapiro-wilk normality test
```{r}
activity_summary %>%
  group_by(sex, genotype) %>%
  shapiro_test(total_act) %>%
  subset(p < 0.05)
```
5 groups not normal

#convert sex and genotype variables to factors
```{r}
activity_summary$genotype <- as.factor(activity_summary$genotype)
activity_summary$sex <- as.factor(activity_summary$sex)
```

#Kruskal test (non-parametric tests for main effects)
```{r}
kruskal.test(total_act ~ sex, data = activity_summary)
kruskal.test(total_act ~ genotype, data = activity_summary)
```
Sex and genotype are significant
Both p-values are references in the manuscript
P=9.0e-12
P=3.2e-09

#subset by genotype and perform Kruskal test comparing sexes, p-values in text
```{r}
dgrp301 <- subset(activity_summary, genotype == "301")
dgrp304 <- subset(activity_summary, genotype == "304")
dgrp786 <- subset(activity_summary, genotype == "786")
dgrp852 <- subset(activity_summary, genotype == "852")

kruskal.test(total_act ~ sex, data = dgrp301)
kruskal.test(total_act ~ sex, data = dgrp304)
kruskal.test(total_act ~ sex, data = dgrp786)
kruskal.test(total_act ~ sex, data = dgrp852)
```
Sex is significant for 301, 786, and 852
The P-value for 852 is referenced in the manuscript text
P=3.4e-06

#outlier and extreme outlier detection
```{r}
activity_summary %>%
  group_by(sex, genotype) %>%
  mutate(
    Q1 = quantile(total_act, 0.25, na.rm = TRUE),
    Q3 = quantile(total_act, 0.75, na.rm = TRUE),
    IQR = Q3 - Q1,
    lower_bound = Q1 - 1.5 * IQR,  # Regular lower bound for outliers
    upper_bound = Q3 + 1.5 * IQR,  # Regular upper bound for outliers
    extreme_lower_bound = Q1 - 3 * IQR,  # Extreme lower bound for extreme outliers
    extreme_upper_bound = Q3 + 3 * IQR,  # Extreme upper bound for extreme outliers
    outlier_flag = total_act < lower_bound | total_act > upper_bound,
    extreme_outlier_flag = total_act < extreme_lower_bound | total_act > extreme_upper_bound
  ) %>%
  ungroup() %>%
  summarize(
    outlier_count = sum(outlier_flag),
    extreme_outlier_count = sum(extreme_outlier_flag)
  )
```
7 outliers and 3 extreme outliers

#remove all the outliers
```{r}
activity_summary_outliers <- activity_summary %>%
  group_by(sex, genotype) %>%
  mutate(
    Q1 = quantile(total_act, 0.25, na.rm = TRUE),
    Q3 = quantile(total_act, 0.75, na.rm = TRUE),
    IQR = Q3 - Q1,
    lower_bound = Q1 - 1.5 * IQR,
    upper_bound = Q3 + 1.5 * IQR,
    outlier_flag = total_act < lower_bound | total_act > upper_bound
  ) %>%
  ungroup() %>%
  filter(outlier_flag == FALSE) %>%
  select(-Q1, -Q3, -IQR, -lower_bound, -upper_bound)
```

#remove only the extreme outliers
```{r}
activity_summary_extremeoutliers <- activity_summary %>%
  group_by(sex, genotype) %>%
  mutate(
    Q1 = quantile(total_act, 0.25, na.rm = TRUE),
    Q3 = quantile(total_act, 0.75, na.rm = TRUE),
    IQR = Q3 - Q1,
    extreme_lower_bound = Q1 - 3 * IQR,
    extreme_upper_bound = Q3 + 3 * IQR,
    extreme_outlier_flag = total_act < extreme_lower_bound | total_act > extreme_upper_bound
  ) %>%
  ungroup() %>%
  filter(!extreme_outlier_flag) %>%
  select(-Q1, -Q3, -IQR, -extreme_lower_bound, -extreme_upper_bound)
```

We realized the ordering of our variables in a regular ANOVA model matters so we are switching to a Type III sums of squares design to deal with these inconsistencies
#mixed-effects model analysis with Type III ANOVA
```{r}
activity_summary_aov <- aov(total_act ~ sex * genotype, data = activity_summary)
Anova(activity_summary_aov, type = 3)

activity_summary_outliers_aov <- aov(total_act ~ sex * genotype, data = activity_summary_outliers)
Anova(activity_summary_outliers_aov, type = 3)

activity_summary_extremeoutliers_aov <- aov(total_act ~ sex * genotype, data = activity_summary_extremeoutliers)
Anova(activity_summary_extremeoutliers_aov, type = 3)
```
Second ANOVA used in Supplementary Table 1

Sex, genotype, and sex:genotype are significant
#now compare models with AIC function for females and males
```{r}
AIC(activity_summary_aov, activity_summary_outliers_aov, activity_summary_extremeoutliers_aov)
```
The dataset with the outliers removed has the lowest/best AIC score

We will run a Q-Q plot and normality test for the best dataset
# Q-Q plot of residuals
```{r}
residuals <- resid(activity_summary_outliers_aov)
qqnorm(residuals)
qqline(residuals, col = "red")
```
#check normality no residuals
```{r}
shapiro.test(residuals)
```
Data is not normal

# Pairwise t-test with multiple corrections
```{r}
pairwise_sex <- pairwise.t.test(activity_summary_outliers$total_act, activity_summary_outliers$sex)
print(pairwise_sex)

pairwise_genotype <- pairwise.t.test(activity_summary_outliers$total_act, activity_summary_outliers$genotype)
print(pairwise_genotype)
```



Analysis of controls at the other timepoints t0, t2, and t3
We want to know if sex, genotype, and sex:genotype remain significant throughout the lifespan.
---
T0 Analysis
#subset data for control groups on day 11
```{r}
genotype_controls0 <- subset(dt, timepoint == "t0")
genotype_controls_t0 <- subset(genotype_controls0, treatment == "c")
```
#aggregate the data by region_id to analyze the total activity
```{r}
activity_summary_t0 <- genotype_controls_t0 %>% 
  group_by(genotype, sex, region_id) %>% 
  summarise(
    mean_act = mean(nd, na.rm = TRUE),
    total_act = sum(nd, na.rm = TRUE),
    .groups = "drop"
  )
```
#shapiro-wilk normality test
```{r}
activity_summary_t0 %>%
  group_by(sex, genotype) %>%
  shapiro_test(total_act) %>%
  subset(p < 0.05)
```
1 group not normal

#convert variables to factor
```{r}
activity_summary_t0$genotype <- as.factor(activity_summary_t0$genotype)
activity_summary_t0$sex <- as.factor(activity_summary_t0$sex)
```

#Kruskal test (non-parametric tests for main effects)
```{r}
kruskal.test(total_act ~ sex, data = activity_summary_t0)
kruskal.test(total_act ~ genotype, data = activity_summary_t0)
```

#outlier and extreme outlier detection
```{r}
activity_summary_t0 %>%
  group_by(sex, genotype) %>%
  mutate(
    Q1 = quantile(total_act, 0.25, na.rm = TRUE),
    Q3 = quantile(total_act, 0.75, na.rm = TRUE),
    IQR = Q3 - Q1,
    lower_bound = Q1 - 1.5 * IQR,  # Regular lower bound for outliers
    upper_bound = Q3 + 1.5 * IQR,  # Regular upper bound for outliers
    extreme_lower_bound = Q1 - 3 * IQR,  # Extreme lower bound for extreme outliers
    extreme_upper_bound = Q3 + 3 * IQR,  # Extreme upper bound for extreme outliers
    outlier_flag = total_act < lower_bound | total_act > upper_bound,
    extreme_outlier_flag = total_act < extreme_lower_bound | total_act > extreme_upper_bound
  ) %>%
  ungroup() %>%
  summarize(
    outlier_count = sum(outlier_flag),
    extreme_outlier_count = sum(extreme_outlier_flag)
  )
```
7 outliers

#remove the outlier
```{r}
activity_summary_t0_outliers <- activity_summary_t0 %>%
  group_by(sex, genotype) %>%
  mutate(
    Q1 = quantile(total_act, 0.25, na.rm = TRUE),
    Q3 = quantile(total_act, 0.75, na.rm = TRUE),
    IQR = Q3 - Q1,
    lower_bound = Q1 - 1.5 * IQR,
    upper_bound = Q3 + 1.5 * IQR,
    outlier_flag = total_act < lower_bound | total_act > upper_bound
  ) %>%
  ungroup() %>%
  filter(outlier_flag == FALSE) %>%
  select(-Q1, -Q3, -IQR, -lower_bound, -upper_bound)
```

#mixed-effects model analysis with Type III ANOVA
```{r}
activity_summary_t0_aov <- aov(total_act ~ sex * genotype, data = activity_summary_t0)
Anova(activity_summary_t0_aov, type = 3)

activity_summary_t0_outliers_aov <- aov(total_act ~ sex * genotype, data = activity_summary_t0_outliers)
Anova(activity_summary_t0_outliers_aov, type = 3)
```
Second ANOVA used in Supplementary Table 1


Sex, genotype, and sex:genotype are significant
#now compare models with AIC function for females and males
```{r}
AIC(activity_summary_t0_aov, activity_summary_t0_outliers_aov)
```
The dataset with the outliers removed has the lowest/best AIC score

We will run a Q-Q plot and normality test for the best dataset
# Q-Q plot of residuals
```{r}
residuals_t0 <- resid(activity_summary_t0_outliers_aov)
qqnorm(residuals_t0)
qqline(residuals_t0, col = "red")
```
#check normality no residuals
```{r}
shapiro.test(residuals_t0)
```
Data is not normal


t0 sex, genotype, and sex:genotype are significant. This is the very first timepoint on day 11.


---
T2 Analysis
---

#subset data for control groups on day 37
```{r}
genotype_controls2 <- subset(dt, timepoint == "t2")
genotype_controls_t2 <- subset(genotype_controls2, treatment == "c")
```

#aggregate the data by region_id to analyze the total activity
```{r}
activity_summary_t2 <- genotype_controls_t2 %>% 
  group_by(genotype, sex, region_id) %>% 
  summarise(
    mean_act = mean(nd, na.rm = TRUE),
    total_act = sum(nd, na.rm = TRUE),
    .groups = "drop"
  )
```

#shapiro-wilk normality test
```{r}
activity_summary_t2 %>%
  group_by(sex, genotype) %>%
  shapiro_test(total_act) %>%
  subset(p < 0.05)
```
3 groups not normal

#convert variables to factor
```{r}
activity_summary_t2$genotype <- as.factor(activity_summary_t2$genotype)
activity_summary_t2$sex <- as.factor(activity_summary_t2$sex)
```

#Kruskal test (non-parametric tests for main effects)
```{r}
kruskal.test(total_act ~ sex, data = activity_summary_t2)
kruskal.test(total_act ~ genotype, data = activity_summary_t2)
```

# Outlier and Extreme Outlier Detection for t2
```{r}
activity_summary_t2 %>%
  group_by(sex, genotype) %>%
  mutate(
    Q1 = quantile(total_act, 0.25, na.rm = TRUE),
    Q3 = quantile(total_act, 0.75, na.rm = TRUE),
    IQR = Q3 - Q1,
    lower_bound = Q1 - 1.5 * IQR,  # Regular lower bound for outliers
    upper_bound = Q3 + 1.5 * IQR,  # Regular upper bound for outliers
    extreme_lower_bound = Q1 - 3 * IQR,  # Extreme lower bound for extreme outliers
    extreme_upper_bound = Q3 + 3 * IQR,  # Extreme upper bound for extreme outliers
    outlier_flag = total_act < lower_bound | total_act > upper_bound,
    extreme_outlier_flag = total_act < extreme_lower_bound | total_act > extreme_upper_bound
  ) %>%
  ungroup() %>%
  summarize(
    outlier_count = sum(outlier_flag),
    extreme_outlier_count = sum(extreme_outlier_flag)
  )
```
4 outliers, 1 extreme

# Removing the Outliers for t2
```{r}
activity_summary_t2_outliers <- activity_summary_t2 %>%
  group_by(sex, genotype) %>%
  mutate(
    Q1 = quantile(total_act, 0.25, na.rm = TRUE),
    Q3 = quantile(total_act, 0.75, na.rm = TRUE),
    IQR = Q3 - Q1,
    lower_bound = Q1 - 1.5 * IQR,
    upper_bound = Q3 + 1.5 * IQR,
    outlier_flag = total_act < lower_bound | total_act > upper_bound
  ) %>%
  ungroup() %>%
  filter(outlier_flag == FALSE) %>%
  select(-Q1, -Q3, -IQR, -lower_bound, -upper_bound)
```

#remove only the extreme outliers
```{r}
activity_summary_t2_extremeoutliers <- activity_summary_t2 %>%
  group_by(sex, genotype) %>%
  mutate(
    Q1 = quantile(total_act, 0.25, na.rm = TRUE),
    Q3 = quantile(total_act, 0.75, na.rm = TRUE),
    IQR = Q3 - Q1,
    extreme_lower_bound = Q1 - 3 * IQR,
    extreme_upper_bound = Q3 + 3 * IQR,
    extreme_outlier_flag = total_act < extreme_lower_bound | total_act > extreme_upper_bound
  ) %>%
  ungroup() %>%
  filter(!extreme_outlier_flag) %>%
  select(-Q1, -Q3, -IQR, -extreme_lower_bound, -extreme_upper_bound)
```

# Mixed-Effects Model Analysis with Type III ANOVA for t2
```{r}
activity_summary_t2_aov <- aov(total_act ~ sex * genotype, data = activity_summary_t2)
Anova(activity_summary_t2_aov, type = 3)

activity_summary_t2_outliers_aov <- aov(total_act ~ sex * genotype, data = activity_summary_t2_outliers)
Anova(activity_summary_t2_outliers_aov, type = 3)

activity_summary_t2_extremeoutliers_aov <- aov(total_act ~ sex * genotype, data = activity_summary_t2_extremeoutliers)
Anova(activity_summary_t2_extremeoutliers_aov, type = 3)
```
Second ANOVA used in Supplementary Table 1


# Comparing Models with AIC for t2
```{r}
AIC(activity_summary_t2_aov, activity_summary_t2_outliers_aov, activity_summary_t2_extremeoutliers_aov)
```
Again outliers data is better

# Q-Q Plot of Residuals for t2
```{r}
residuals_t2 <- resid(activity_summary_t2_outliers_aov)
qqnorm(residuals_t2)
qqline(residuals_t2, col = "red")
```
# Check Normality of Residuals for t2
```{r}
shapiro.test(residuals_t2)
```


Sex and sex:genotype are significant but we lost genotype according to the type 3 anova. Kruskal test still claims genotype is significant.


---
T3 Analysis
---
#subset data for control groups on day 51
```{r}
genotype_controls3 <- subset(dt, timepoint == "t3")
genotype_controls_t3 <- subset(genotype_controls3, treatment == "c")
```
#aggregate the data by region_id to analyze the total activity
```{r}
activity_summary_t3 <- genotype_controls_t3 %>% 
  group_by(genotype, sex, region_id) %>% 
  summarise(
    mean_act = mean(nd, na.rm = TRUE),
    total_act = sum(nd, na.rm = TRUE),
    .groups = "drop"
  )
```
#shapiro-wilk normality test
```{r}
activity_summary_t3 %>%
  group_by(sex, genotype) %>%
  shapiro_test(total_act) %>%
  subset(p < 0.05)
```
5 groups not normal

#convert variables to factors
```{r}
activity_summary_t3$genotype <- as.factor(activity_summary_t3$genotype)
activity_summary_t3$sex <- as.factor(activity_summary_t3$sex)
```
#Kruskal test (non-parametric tests for main effects)
```{r}
kruskal.test(total_act ~ sex, data = activity_summary_t3)
kruskal.test(total_act ~ genotype, data = activity_summary_t3)
```

# Outlier and Extreme Outlier Detection for t3
```{r}
activity_summary_t3 %>%
  group_by(sex, genotype) %>%
  mutate(
    Q1 = quantile(total_act, 0.25, na.rm = TRUE),
    Q3 = quantile(total_act, 0.75, na.rm = TRUE),
    IQR = Q3 - Q1,
    lower_bound = Q1 - 1.5 * IQR,  # Regular lower bound for outliers
    upper_bound = Q3 + 1.5 * IQR,  # Regular upper bound for outliers
    extreme_lower_bound = Q1 - 3 * IQR,  # Extreme lower bound for extreme outliers
    extreme_upper_bound = Q3 + 3 * IQR,  # Extreme upper bound for extreme outliers
    outlier_flag = total_act < lower_bound | total_act > upper_bound,
    extreme_outlier_flag = total_act < extreme_lower_bound | total_act > extreme_upper_bound
  ) %>%
  ungroup() %>%
  summarize(
    outlier_count = sum(outlier_flag),
    extreme_outlier_count = sum(extreme_outlier_flag)
  )
```
8 outliers, 4 extreme outliers

# Removing the Outliers for t3
```{r}
activity_summary_t3_outliers <- activity_summary_t3 %>%
  group_by(sex, genotype) %>%
  mutate(
    Q1 = quantile(total_act, 0.25, na.rm = TRUE),
    Q3 = quantile(total_act, 0.75, na.rm = TRUE),
    IQR = Q3 - Q1,
    lower_bound = Q1 - 1.5 * IQR,
    upper_bound = Q3 + 1.5 * IQR,
    outlier_flag = total_act < lower_bound | total_act > upper_bound
  ) %>%
  ungroup() %>%
  filter(outlier_flag == FALSE) %>%
  select(-Q1, -Q3, -IQR, -lower_bound, -upper_bound)
```
# Remove only the extreme outliers for t3
```{r}
activity_summary_t3_extremeoutliers <- activity_summary_t3 %>%
  group_by(sex, genotype) %>%
  mutate(
    Q1 = quantile(total_act, 0.25, na.rm = TRUE),
    Q3 = quantile(total_act, 0.75, na.rm = TRUE),
    IQR = Q3 - Q1,
    extreme_lower_bound = Q1 - 3 * IQR,
    extreme_upper_bound = Q3 + 3 * IQR,
    extreme_outlier_flag = total_act < extreme_lower_bound | total_act > extreme_upper_bound
  ) %>%
  ungroup() %>%
  filter(!extreme_outlier_flag) %>%
  select(-Q1, -Q3, -IQR, -extreme_lower_bound, -extreme_upper_bound)
```

# Mixed-Effects Model Analysis with Type III ANOVA for t3
```{r}
activity_summary_t3_aov <- aov(total_act ~ sex * genotype, data = activity_summary_t3)
Anova(activity_summary_t3_aov, type = 3)

activity_summary_t3_outliers_aov <- aov(total_act ~ sex * genotype, data = activity_summary_t3_outliers)
Anova(activity_summary_t3_outliers_aov, type = 3)

activity_summary_t3_extremeoutliers_aov <- aov(total_act ~ sex * genotype, data = activity_summary_t3_extremeoutliers)
Anova(activity_summary_t3_extremeoutliers_aov, type = 3)
```
Second ANOVA used in Supplementary Table 1


# Comparing Models with AIC for t3
```{r}
AIC(activity_summary_t3_aov, activity_summary_t3_outliers_aov, activity_summary_t3_extremeoutliers_aov)
```
Again outliers data is better

# Q-Q Plot of Residuals for t3
```{r}
residuals_t3 <- resid(activity_summary_t3_outliers_aov)
qqnorm(residuals_t3)
qqline(residuals_t3, col = "red")
```
# Check Normality of Residuals for t3
```{r}
shapiro.test(residuals_t3)
```
Data is not normal

Sex and sex:genotype are significant, but once again we have lost genotype from the type 3 ANOVA. Genotype was significant according to Kruskal Wallis test.

