---
title: "Supplemental Figure 1"
output: html_document
date: "2025-05-30"
---

#load the library
```{r}
library(readxl)
library(car)
library(lme4)
library(reshape2)
library(ggplot2)
library(reshape)
library(dplyr)
library(here)
library(ARTool)
library(rstatix)
library(damr)
library(ggetho)
```
With "here" we do not need to set a working directory. Keep script and data in same folder as the Data subfolder (has .txt files) and the metadata.csv - It should already be setup and ready to use in the Data 1.zip.

#clear workspace
```{r}
rm(list = ls())
```
#check working directory
```{r}
getwd()
```

First, look at 786 female total activity data on boxplots
Load data frames for DGRP786 DAM data
#set file name
```{r}
DAM_786 <- "DAM_786_2023_11_06.xlsx"
```

Steps to load and organize all the data for the t0 timepoint
#load the normalized data into a dataframe for each timepoint/treatment
```{r}
  t0_c_786 <- read_excel(DAM_786, sheet = "t0_c_norm")
  t0_5d_786 <- read_excel(DAM_786, sheet = "t0_5d_norm")
```
#load the legend specific to the timepoint t0
```{r}
  legend_786 <- read_excel(DAM_786, sheet = "legend_t0")
  legend_786$chan_no <- as.factor(legend_786$chan_no)
```
#break the legend out by treatment in order to do a left join 
```{r}
  legend_c_786 <- subset(legend_786, treatment == "c")
  legend_5d_786 <- subset(legend_786, treatment== "5d")
```
#convert to data frame
```{r}
  t0_c_786 <- as.data.frame(t0_c_786)
  t0_5d_786 <- as.data.frame(t0_5d_786)
```
#use the melt function to convert to long format
```{r}
  t0_c_786 <- melt(t0_c_786, id.vars = c("index", "date", "time"), 
                   measure.vars = c("1", "2", "3", "4", "5", "6", "7", "8",
                                    "9", "10", "11", "12", "13", "14", "15",
                                    "16", "17", "18", "19", "20", "21", "22",
                                    "23", "24", "25", "26", "27", "28", "29",
                                    "30", "31", "32"))
  t0_5d_786 <- melt(t0_5d_786, id.vars = c("index", "date", "time"), 
                    measure.vars = c("1", "2", "3", "4", "5", "6", "7", "8",
                                     "9", "10", "11", "12", "13", "14", "15",
                                     "16", "17", "18", "19", "20", "21", "22",
                                     "23", "24", "25", "26", "27", "28", "29",
                                     "30", "31", "32"))
```
#sanity check 1
```{r}
  summary(t0_c_786)
  summary(t0_5d_786)
```
#after data is in long format, sum the 5-minute intervals into one daily total
```{r}
  t0_c_786 <- aggregate(t0_c_786$value, by=list(chan_no=t0_c_786$variable), FUN=sum)
  t0_5d_786 <- aggregate(t0_5d_786$value, by=list(chan_no=t0_5d_786$variable), FUN=sum)
```
#rename columns from variable and value
```{r}
  colnames(t0_c_786) <- c("chan_no", "daily_activity")
  colnames(t0_5d_786) <- c("chan_no", "daily_activity")
```
#unblind using the legends- specific to treatment- on the vial_no value
```{r}
  t0_c_786 <- left_join(t0_c_786, legend_c_786, by = "chan_no")
  t0_5d_786 <- left_join(t0_5d_786, legend_5d_786, by = "chan_no")
```
#get rid of unneeded columns
```{r}
  keeps <- c("chan_no", "daily_activity", "treatment", "genotype", "sex", "age")
  t0_c_786 <- t0_c_786[keeps]
  t0_5d_786 <- t0_5d_786[keeps]
```
#combine the treatments into one master per timepoint using rbind
```{r}
  master_t0_786 <- rbind(t0_c_786, t0_5d_786)
```
#sanity check 2
```{r}
  summary(master_t0_786)

  hist(master_t0_786$daily_activity)
  abline(v = mean(master_t0_786$daily_activity), col="red", lwd=3)
```
---
Steps to load and organize all the data for the t1 timepoint
#load the normalized data into a dataframe for each timepoint/treatment
```{r}
  t1_c_786 <- read_excel(DAM_786, sheet = "t1_c_norm")
  t1_5d_786 <- read_excel(DAM_786, sheet = "t1_5d_norm")
  t1_20d_786 <- read_excel(DAM_786, sheet = "t1_20d_norm")
```
#load the legend specific to the timepoint
```{r}
  legend_786 <- read_excel(DAM_786, sheet = "legend_t1")
  legend_786$chan_no <- as.factor(legend_786$chan_no)
```
#break the legend out by treatment in order to do a left join
```{r}
  legend_c_786 <- subset(legend_786, treatment == "c")
  legend_5d_786 <- subset(legend_786, treatment== "5d")
  legend_20d_786 <- subset(legend_786, treatment== "20d")
```
#convert to data frame
```{r}
  t1_c_786 <- as.data.frame(t1_c_786)
  t1_5d_786 <- as.data.frame(t1_5d_786)
  t1_20d_786 <- as.data.frame(t1_20d_786)
```
#use the melt function to convert to long format
```{r}
  t1_c_786 <- melt(t1_c_786, id.vars = c("index", "date", "time"), 
                   measure.vars = c("1", "2", "3", "4", "5", "6", "7", "8",
                                    "9", "10", "11", "12", "13", "14", "15",
                                    "16", "17", "18", "19", "20", "21", "22",
                                    "23", "24", "25", "26", "27", "28", "29",
                                    "30", "31", "32"))
  t1_5d_786 <- melt(t1_5d_786, id.vars = c("index", "date", "time"), 
                    measure.vars = c("1", "2", "3", "4", "5", "6", "7", "8",
                                     "9", "10", "11", "12", "13", "14", "15",
                                     "16", "17", "18", "19", "20", "21", "22",
                                     "23", "24", "25", "26", "27", "28", "29",
                                     "30", "31", "32"))
  t1_20d_786 <- melt(t1_20d_786, id.vars = c("index", "date", "time"), 
                     measure.vars = c("1", "2", "3", "4", "5", "6", "7", "8",
                                      "9", "11", "12", "13", "14", "15",
                                      "16", "17", "18", "19", "20", "21", "22",
                                      "23", "24", "25", "26", "27", "28", "29",
                                      "30", "31", "32"))
```
#sanity check 1
```{r}
 summary(t1_c_786)
  summary(t1_5d_786)
  summary(t1_20d_786)
```
#after data is in long format, sum the 5-minute intervals into one daily total
```{r}
  t1_c_786 <- aggregate(t1_c_786$value, by=list(chan_no=t1_c_786$variable), FUN=sum)
  t1_5d_786 <- aggregate(t1_5d_786$value, by=list(chan_no=t1_5d_786$variable), FUN=sum)
  t1_20d_786 <- aggregate(t1_20d_786$value, by=list(chan_no=t1_20d_786$variable), FUN=sum)
```
#rename columns from variable and value
```{r}
  colnames(t1_c_786) <- c("chan_no", "daily_activity")
  colnames(t1_5d_786) <- c("chan_no", "daily_activity")
  colnames(t1_20d_786) <- c("chan_no", "daily_activity")
```
#unblind using the legends- specific to treatment- on the vial_no value
```{r}
  t1_c_786 <- left_join(t1_c_786, legend_c_786, by = "chan_no")
  t1_5d_786 <- left_join(t1_5d_786, legend_5d_786, by = "chan_no")
  t1_20d_786 <- left_join(t1_20d_786, legend_20d_786, by = "chan_no")
```
#get rid of unneeded columns
```{r}
  keeps <- c("chan_no", "daily_activity", "treatment", "genotype", "sex", "age")
  t1_c_786 <- t1_c_786[keeps]
  t1_5d_786 <- t1_5d_786[keeps]
  t1_20d_786 <- t1_20d_786[keeps]
```
#combine the treatments into one master per timepoint using rbind
```{r}
  master_t1_786 <- rbind(t1_c_786, t1_5d_786, t1_20d_786)
```
#sanity check 2
```{r}
  summary(master_t1_786)

  hist(master_t1_786$daily_activity)
  abline(v = mean(master_t1_786$daily_activity), col="red", lwd=3)
```
---
Steps to load and organize all the data for the t2 timepoint
#load the normalized data into a dataframe for each timepoint/treatment
```{r}
  t2_c_786 <- read_excel(DAM_786, sheet = "t2_c_norm")
  t2_5d_786 <- read_excel(DAM_786, sheet = "t2_5d_norm")
  t2_20d_786 <- read_excel(DAM_786, sheet = "t2_20d_norm")
```
#load the legend specific to the timepoint t1_t2
```{r}
  legend_t1_t2_786 <- read_excel(DAM_786, sheet = "legend_t2")
  legend_t1_t2_786$chan_no <- as.factor(legend_t1_t2_786$chan_no)
```
#break the legend out by treatment in order to do a left join
```{r}
  legend_c_786 <- subset(legend_t1_t2_786, treatment == "c")
  legend_5d_786 <- subset(legend_t1_t2_786, treatment== "5d")
  legend_20d_786 <- subset(legend_t1_t2_786, treatment== "20d")
```
#convert to data frame
```{r}
  t2_c_786 <- as.data.frame(t2_c_786)
  t2_5d_786 <- as.data.frame(t2_5d_786)
  t2_20d_786 <- as.data.frame(t2_20d_786)
```
#use the melt function to convert to long format
```{r}
  t2_c_786 <- melt(t2_c_786, id.vars = c("index", "date", "time"), 
                   measure.vars = c("1", "2", "3", "4", "5", "6", "7", "8",
                                    "9", "10", "11", "12", "13", "14", "15",
                                    "16", "17", "18", "19", "20", "21", "22",
                                    "23", "24", "25", "26", "27", "28", "29",
                                    "30", "31", "32"))
  t2_5d_786 <- melt(t2_5d_786, id.vars = c("index", "date", "time"), 
                    measure.vars = c("1", "2", "3", "4", "5", "6", "7", "8",
                                     "9", "10", "11", "12", "13", "14", "15",
                                     "16", "17", "18", "19", "20", "21", "22",
                                     "23", "24", "25", "26", "27", "28", "29",
                                     "30", "31", "32"))
  t2_20d_786 <- melt(t2_20d_786, id.vars = c("index", "date", "time"), 
                     measure.vars = c("1", "2", "3", "4", "5", "6", "7", "8",
                                      "9", "11", "12", "13", "14", "15",
                                      "16", "17", "18", "19", "20", "21", "22",
                                      "23", "24", "25", "26", "27", "28", "29",
                                      "30", "31", "32"))
```
#sanity check 1
```{r}
  summary(t2_c_786)
  summary(t2_5d_786)
  summary(t2_20d_786)
```
#after data is in long format, sum the 5-minute intervals into one daily total
```{r}
  t2_c_786 <- aggregate(t2_c_786$value, by=list(chan_no=t2_c_786$variable), FUN=sum)
  t2_5d_786 <- aggregate(t2_5d_786$value, by=list(chan_no=t2_5d_786$variable), FUN=sum)
  t2_20d_786 <- aggregate(t2_20d_786$value, by=list(chan_no=t2_20d_786$variable), FUN=sum)
```
#rename columns from variable and value
```{r}
  colnames(t2_c_786) <- c("chan_no", "daily_activity")
  colnames(t2_5d_786) <- c("chan_no", "daily_activity")
  colnames(t2_20d_786) <- c("chan_no", "daily_activity")
```
#unblind using the legends- specific to treatment- on the vial_no value
```{r}
  t2_c_786 <- left_join(t2_c_786, legend_c_786, by = "chan_no")
  t2_5d_786 <- left_join(t2_5d_786, legend_5d_786, by = "chan_no")
  t2_20d_786 <- left_join(t2_20d_786, legend_20d_786, by = "chan_no")
```
#get rid of unneeded columns
```{r}
  keeps <- c("chan_no", "daily_activity", "treatment", "genotype", "sex", "age")
  t2_c_786 <- t2_c_786[keeps]
  t2_5d_786 <- t2_5d_786[keeps]
  t2_20d_786 <- t2_20d_786[keeps]
```
#combine the treatments into one master per timepoint using rbind
```{r}
  master_t2_786 <- rbind(t2_c_786, t2_5d_786, t2_20d_786)
```
#sanity check 2
```{r}
  summary(t2_20d_786)

  hist(t2_20d_786$daily_activity)
  abline(v = mean(t2_20d_786$daily_activity), col="red", lwd=3)
```
---
Steps to load and organize all the data for the t3 timepoint
#load the normalized data into a dataframe for each timepoint/treatment
```{r}
  t3_c_786 <- read_excel(DAM_786, sheet = "t3_c_norm")
  t3_5d_786 <- read_excel(DAM_786, sheet = "t3_5d_norm")
  t3_20d_786 <- read_excel(DAM_786, sheet = "t3_20d_norm")
```
#load the legend specific to the timepoint t3
```{r}
  legend_t3_786 <- read_excel(DAM_786, sheet = "legend_t3")
  legend_t3_786$chan_no <- as.factor(legend_t3_786$chan_no)
```
#break the legend out by treatment in order to do a left join
```{r}
  legend_c_786 <- subset(legend_t3_786, treatment == "c")
  legend_5d_786 <- subset(legend_t3_786, treatment== "5d")
  legend_20d_786 <- subset(legend_t3_786, treatment== "20d")
```
#convert to data frame
```{r}
  t3_c_786 <- as.data.frame(t3_c_786)
  t3_5d_786 <- as.data.frame(t3_5d_786)
  t3_20d_786 <- as.data.frame(t3_20d_786)
```
#use the melt function to convert to long format
```{r}
  t3_c_786 <- melt(t3_c_786, id.vars = c("index", "date", "time"), 
                   measure.vars = c("1", "2", "3", "4", "5", "6", "7", "8",
                                    "9", "10", "11", "12", "13", "14", "15",
                                    "16", "17", "18", "19", "20", "21", "22",
                                    "23", "24", "25", "26", "27", "28", "29",
                                    "30", "31", "32"))
  t3_5d_786 <- melt(t3_5d_786, id.vars = c("index", "date", "time"), 
                    measure.vars = c("1", "2", "3", "4", "5", "6", "7", "8",
                                     "9", "10", "11", "12", "13", "14", "15",
                                     "16", "17", "18", "19", "20", "21", "22",
                                     "23", "24", "25", "26", "27", "28", "29",
                                     "30", "31", "32"))
  t3_20d_786 <- melt(t3_20d_786, id.vars = c("index", "date", "time"), 
                     measure.vars = c("1", "2", "3", "4", "5", "6", "7", "8",
                                      "9", "11", "12", "13", "14", "15",
                                      "16", "17", "18", "19", "20", "21", "22",
                                      "23", "24", "25", "26", "27", "28", "29",
                                      "30", "31", "32"))
```
#sanity check 1
```{r}
  summary(t3_c_786)
  summary(t3_5d_786)
  summary(t3_20d_786)
```
#after data is in long format, sum the 5-minute intervals into one daily total
```{r}
  t3_c_786 <- aggregate(t3_c_786$value, by=list(chan_no=t3_c_786$variable), FUN=sum)
  t3_5d_786 <- aggregate(t3_5d_786$value, by=list(chan_no=t3_5d_786$variable), FUN=sum)
  t3_20d_786 <- aggregate(t3_20d_786$value, by=list(chan_no=t3_20d_786$variable), FUN=sum)
```
#rename columns from variable and value
```{r}
  colnames(t3_c_786) <- c("chan_no", "daily_activity")
  colnames(t3_5d_786) <- c("chan_no", "daily_activity") 
  colnames(t3_20d_786) <- c("chan_no", "daily_activity")
```
#unblind using the legends- specific to treatment- on the vial_no value
```{r}
  t3_c_786 <- left_join(t3_c_786, legend_c_786, by = "chan_no")
  t3_5d_786 <- left_join(t3_5d_786, legend_5d_786, by = "chan_no")
  t3_20d_786 <- left_join(t3_20d_786, legend_20d_786, by = "chan_no")
```
#get rid of unneeded columns
```{r}
  keeps <- c("chan_no", "daily_activity", "treatment", "genotype", "sex", "age")
  t3_c_786 <- t3_c_786[keeps]
  t3_5d_786 <- t3_5d_786[keeps]
  t3_20d_786 <- t3_20d_786[keeps]
```
#combine the treatments into one master per timepoint using rbind
```{r}
  master_t3_786 <- rbind(t3_c_786, t3_5d_786, t3_20d_786)
```
#sanity check 2
```{r}
  summary(t3_c_786)

  hist(master_t3_786$daily_activity)
  abline(v = mean(master_t3_786$daily_activity), col="red", lwd=3)
```
#combine all of 786
```{r}
  master_786 <- rbind(master_t0_786, master_t1_786, master_t2_786, master_t3_786)
```
#remove outliers
```{r}
master_786_outliers <- master_786 %>%
  group_by(treatment) %>%
  mutate(
    Q1 = quantile(daily_activity, 0.25, na.rm = TRUE),
    Q3 = quantile(daily_activity, 0.75, na.rm = TRUE),
    IQR = Q3 - Q1,
    lower_bound = Q1 - 1.5 * IQR,
    upper_bound = Q3 + 1.5 * IQR,
    outlier_flag = daily_activity < lower_bound | daily_activity > upper_bound
  ) %>%
  ungroup() %>%
  filter(outlier_flag == FALSE) %>%
  select(-Q1, -Q3, -IQR, -lower_bound, -upper_bound)
```
#pull out females for the figure
```{r}
master_786_females <- subset(master_786_outliers, sex == "F")
```

#plot DGRP786 females
```{r}
master_786_females$treatment <- factor(master_786_females$treatment,
                                        levels = c("c", "5d", "20d"))
boxplot_786_females <- ggplot(data = master_786_females %>%
         mutate(treatment = factor(treatment, levels = c("c", "5d", "20d"))),
       aes(x = sex, y = daily_activity)) +
  geom_boxplot(aes(fill = "white", color = treatment), outlier.shape = NA, size=0.75) +
  geom_jitter(aes(x = sex, y = daily_activity, color = treatment), 
              alpha = 0.6, 
              position = position_jitterdodge(jitter.width = 0.2, jitter.height = 0), 
              size = 1) + 
  facet_grid(~age) +
  scale_fill_manual(values = "white") + 
  scale_color_manual(values = c("#FF4433","#EFC000FF","#0073C2FF"),
                     labels = c("Control", "5-day", "20-day")) +
  labs(x = "Sex", y = "Activity per fly", fill = "Treatment", color = "Treatment") +
  theme_bw() +
  theme(strip.background = element_rect(fill = "white"), 
        strip.text = element_text(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) 

boxplot_786_females

```



Second, look at 786 female activity vs time data on line plot.
#store path variable
```{r}
DATA_DIR <- "Data"
```
#check all files are present
```{r}
list.files(DATA_DIR, pattern="*.txt|*.csv")
```
#Define metavariables for the corresponding legend within each genotype
```{r}
data <- fread("metadata.csv")
data
```
#Link meta data based on genotype
```{r}
metadata <- link_dam_metadata(data, result_dir = DATA_DIR)
metadata
```
#Load into a behavr structure
Exclude mated animals not under "OK" status, find and load matching data
```{r}
dt <- load_dam(metadata[status == "OK"])
summary(dt)
```
#Rejoin data and divide by counts
```{r}
dt <- rejoin(dt)
dt[, "nd"] <- dt[, "activity"] / dt[, "counts"]
```
#Turn back into behavr table
```{r}
dt <- behavr(dt, metadata [status == "OK"])
```
#Make 0 the start 7AM lights on time point
```{r}
dt[,t := t - hours(xmv(baseline_time))]
```
Create a graph of day 26 activity v time with control, 5d and 20d groups. This will be used to make Figure 3, and boxes will be added around the times of interest: 6-8AM, 12-2PM, and 6-8PM in Adobe illustrator.
# Subset data from day 26
```{r}
dt_t1 <- subset(dt, timepoint == "t1")
```

# Shift the time for genotypes 786 and 304 by -3600 seconds
DGRP786 and DGRP304 were the only two genotypes impacted by daylights savings. So set them back by 1hr. In the metadata sheet I already set the 0 with the daylights savings accounted for so instead of 7=0 it was 6=0.
```{r}
dt_t1$t <- ifelse(dt_t1$genotype %in% c(786, 304), dt_t1$t - 3600, dt_t1$t)
```
#subset out day 26 data for dgrp786
```{r}
dt_t1_786 <- subset(dt_t1, genotype == "786")
```
#subset out females for dgrp786
```{r}
dt_t1_786_females <- subset(dt_t1_786, sex == "F")
```

#graph 786
```{r}
graph_786_females <- ggetho(dt_t1_786_females, aes(x = t, y = nd, colour = treatment)) + 
  facet_grid(sex ~ genotype) + 
  labs(y = "Activity per fly", x = "Time (h)") + 
  theme_bw() +
  scale_color_manual(values = c("#0073C2FF", "#EFC000FF", "#FF4433"), name = "Treatment") +
  scale_fill_manual(values = c("#0073C2FF", "#EFC000FF", "#FF4433"), name = "Treatment") +
  theme(strip.background = element_rect(fill = "white")) + 
  theme(strip.text = element_text(colour = "black")) +
  annotate('rect', xmin = -25200, xmax = 0, ymin = 0, ymax = 0.145, alpha = 0.5, fill = "darkgrey") +
  annotate('rect', xmin = 43200, xmax = 59500, ymin = 0, ymax = 0.145, alpha = 0.5, fill = "darkgrey") +
  annotate('rect', xmin = -3600, xmax = 3600, ymin = 0, ymax = 0.145, alpha = 0, color = "black", linewidth = 1) +  # 6-8 AM
  annotate('rect', xmin = 18000, xmax = 25200, ymin = 0, ymax = 0.145, alpha = 0, color = "black", linewidth = 1) +  # 12-2 PM
  annotate('rect', xmin = 39600, xmax = 46800, ymin = 0, ymax = 0.145, alpha = 0, color = "black", linewidth = 1) +  # 6-8 PM
  stat_pop_etho() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = "none")

graph_786_females
```





